---
title: "Provincial Choropleth Map"
author: "Julia Gorospe"
date: 2026-09-21
---

```{r setup}
#| message: false

# load packages
library(tidyverse)
library(gt)
#library(sf)
```

## Dataset
source: Natural Earth
accreditation: Made with Natural Earth. Free vector and raster map data @ naturalearthdata.com.
license: Creative Commons CC0 Public Domain Dedication

```{r load data}
#| warning: false

#provinces <- read_sf()

```

## What is a Choropleth Map
Each region is represented as a polygon with boundaries representing the separation between different areas. Zonal statistics may be calculated that correspond with each region. The values can be mapped to a colour or a colour saturation level to create intuitive 


## Visualizing Provinical Summary Statistics
Choropleth maps can be generated in R with several tools including ggplot. Here the provinces are plotted on the map with colours assigned based on population.

```{r map}
world_data <- ggplot2::map_data("world")
pop_data <- gt::countrypops
head(world_data)
ggplot(world_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", color = "white") +
  theme_minimal()
# ggplot(my_sf) +
#   geom_sf(fill = "white", color = "black", linewidth = 0.3) +
#   theme_void()
```