---
title: "Global Energy Imports"
author: "Julia Gorospe"
date: 2026-09-25
---

```{r}
#| label: setup
#| message: false

# load packages
library(tidyverse)

```

## Dataset
The World Bank compiles time series data for a variety of society measures by administrative regions. Here trends in energy import are visualized through a set of maps.

**Indicator description from World Bank**
_Net energy imports are estimated as gross imports less gross exports, both measured in tons of oil equivalents (toe). A negative value indicates that the country is a net exporter. Energy use refers to use of primary energy before transformation to other end-use fuels, which is equal to indigenous production plus imports and stock changes, minus exports and fuels supplied to ships and aircraft engaged in international transport._

source: [World Development Indicators](https://databank.worldbank.org/source/world-development-indicators) Energy imports, net (% of energy use) - EG.IMP.CONS.ZS     
license: Creative Commons Attribution 4.0

A simple global map with country boundaries is available from Natural Earth through the `ggplot2` package.  
source: [Natural Earth](https://www.naturalearthdata.com/downloads/)  
license: Creative Commons CC0 Public Domain Dedication  

```{r}
#| warning: false

# load WDI dataset
energy_raw <- read_csv("energy_imports.csv") |> 
    mutate(`Country Name` = case_when(
      `Country Name` == "Russian Federation" ~ "Russia",
      `Country Name` == "United States" ~ "USA",
      TRUE ~ `Country Name`)
    )

# load world data from ggplot2 package
world_data <- ggplot2::map_data("world")
```

## Data Wrangling
The energy import data is too bulky to visualize meaningfully and needs to be subset to provide a usable dataset. To create the maps, the energy import data must be merged with the geographic data.

```{r}
energy_subset <- energy_raw |> 
  filter(
    Time == "2020",
    `Energy imports, net (% of energy use) [EG.IMP.CONS.ZS]` != ".."
  ) |> 
  mutate(energy_import = as.numeric(`Energy imports, net (% of energy use) [EG.IMP.CONS.ZS]`)) |> 
  filter(energy_import >= -150 & energy_import <=150) |> 
  full_join(world_data, by = join_by(`Country Name` == region))
```


## What is a Choropleth Map
Each region is represented as a polygon with boundaries representing the separation between different areas. Zonal statistics may be calculated that correspond with each region. The values can be mapped to a colour or a colour saturation level to create intuitive 


## Visualizing Country Level Statistics
Choropleth maps can be generated in R with several tools including ggplot. Here countries are plotted on the map with colours assigned based on net energy import percentage.

```{r}
#| label: map
#| fig-width: 8
#| fig-height: 4.5
#| fig-cap: "% of Energy Imported by Country in 2020."

ggplot(energy_subset, aes(x = long, y = lat, group = group, fill = energy_import)) +
  geom_polygon(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "lightgrey", high = "yellow", midpoint = 0) +
  theme_minimal()
```

accreditation: Made with Natural Earth. Free vector and raster map data @ naturalearthdata.com. Data from World Bank - World Development Indicators (2026).